SPARQL on the Open, Decentralised Web

### 2016-06-23 dissertation/introduction.tex
more on HTTP
query planning

Note that all the IRIs uses the HTTP schema. The Hypertext Transfer
Protocol (HTTP) is the main application protocol used on the Internet
for the Web, and is also the protocol used most for the Semantic
Web. It is a request-response protocol in a client-server computing
model, but also accommodates for intermediate proxies.

@@

\begin{verbatim}
?subject a ?class .
\end{verbatim}
This triple pattern will match the two triples 
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
from the above example. Frameworks in popular programming languages
commonly have a method that implements this.

@@

>>>>

Note that all the IRIs uses the HTTP schema. The Hypertext Transfer
Protocol (HTTP) is the main application protocol used on the Internet
for the Web, and is also the protocol used most for the Semantic
Web. It is a request-response protocol in a client-server computing
model, but also accommodates for intermediate proxies. HTTP messages
consist of a header, typically with metadata, and a body, which is the
content itself. The standard defines several headers that can be used
to control caching on either the client, the server or an intermediate
proxy.

@@

\begin{verbatim}
?subject a ?class .
\end{verbatim}
This triple pattern will match the two triples 
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
from the above example (prefixes omitted for brevity). Frameworks in
popular programming languages commonly have a method that implements
this.

@@

To execute a query, a SPARQL engine would typically parse the query to
create a tree of algebra objects (in the case where an object oriented
programming language is used) in the process. Then, a query planner
component would traverse the algebra tree, and for each algebra object
or branch of algebra objects, the query planner finds various ways to
access the data in the underlying store or execute other
operations. In many programming frameworks, the only way to access
data is through matching a triple pattern. If that is the case and
the incoming query consists of just a BGP, the query planner's task
would simply be to match all the triple patterns and join their
results, and finally apply any constraining filters.

However, even though the result does not depend on the order of
execution, the time it takes to find the result often does. For
example, consider the following BGP:
\begin{verbatim}
?subject foaf:name ?name ;
         foaf:workplaceHomepage <http://example.com/> ;
         foaf:made ?contribution .
\end{verbatim}
In most cases, if the triple pattern with the \rdfterm{foaf:name}
predicate is evaluated first, it is likely to match many resources,
since nearly everything may have a name, and therefore, it does not
reduce the number of statements that needs to be searched when
matching the next triple pattern by much. Instead, the
\rdfterm{foaf:workplaceHomepage} may be matched first, as we may guess
the fact that the object is given will be more restrictive, but that
too fails if everyone in the database works for the same organisation
and has used this predicate. It is the query planner's task to choose
between equivalent plans to ensure that the most efficient evaluation
is chosen. As is clear from this example, it is often insufficient to
rely on such heuristics, and therefore, it is common to rely on a
statistical digest in the query planner. Still, the task of query
planning remains a very complex one, and often, insufficient
statistics is available, so that planning is done using heuristics or
based on assumptions that do not hold.

Indeed, detailed knowledge of the data, such as a comprehensive
statistical digest may be available close to the database, but is not
exposed in ways that higher levels in the technology stack can make
use of. In our case, this is especially true for a caching proxy.
